### 2024-06-09 HickernellFredTutorial.tex
Updates from Overleaf

\begin{talk}
  {Low Discrepancy Sequences for Kernel Density Estimation}% [1] talk title
  {Fred J. Hickernell}% [2] speaker name
  {Department of Applied Mathematics and Center for Interdisciplinary Scientific Computation, Illinois Institute of Technology}% [3] affiliations
  {hickernell@iit.edu}% [4] email
  {}% [5] coauthors
  {Universality in QMC and related algorithms}% [6] special session. Leave this field empty for contributed talks. 
				% Insert the title of the special session if you were invited to give a talk in a special session.

Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, are well known to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.  The use of LD sequences for density estimation has been studied recently %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a} 
[1--3].

This talk explores further the use of LD sequences for kernel density estimation, but from a deterministic perspective.  Given a random variable $Y = f(\boldsymbol{X})$, where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$, with a probability density $\varrho$, and kernel density estimate, $\varrho$ as

\medskip

\begin{enumerate}
	\item[{[1]}]
{Ben Abdellah}, A., L'Ecuyer, P., Owen, A.~B., and Puchhammer, F. (2021).
\newblock Density estimation by randomized quasi-{M}onte {C}arlo.
\newblock {\em SIAM/ASA J.\ Uncertain.\ Quantif.}, 9(280-301).

\item[{[2]}]
L'Ecuyer, P. and Puchhammer, F. (2022).
\newblock Density estimation by Monte Carlo and quasi-Monte Carlo.
\newblock In Keller, A., editor, {\em {M}onte {C}arlo and Quasi-{M}onte {C}arlo Methods: {MCQMC}, {O}xford, England, {A}ugust 2020}, Springer Proceedings in Mathematics and Statistics. Springer, Cham.

\item[{[3]}]
L'Ecuyer, P., Puchhammer, F., and {Ben Abdellah}, A. (2022).
\newblock Monte Carlo and quasi{\textendash}{M}onte {C}arlo density estimation via conditioning.
\newblock {\em INFORMS J.\ Comput.}, 34(3):1729--1748.
\end{enumerate}

@@

%\bibliographystyle{apalike}
%\bibliography{FJH23,FJHown23}

>>>>

\begin{talk}
  {Quasi-Monte Carlo Methods:  What, Why, and How?}% [1] talk title
  {Fred J. Hickernell}% [2] speaker name
  {Department of Applied Mathematics and Center for Interdisciplinary Scientific Computation, Illinois Institute of Technology}% [3] affiliations
  {hickernell@iit.edu}% [4] email
  {}% [5] coauthors
  {}% [6] special session. Leave this field empty for contributed talks. 
				% Insert the title of the special session if you were invited to give a talk in a special session.

Many problems in  quantitative finance, uncertainty quantification, and other areas can be formulated as computing $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ are generated by numerical simulation. The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=1}^n f(\boldsymbol{X}_i)$.  Computing $\mu$ is equivalent to computing a $d$-dimensional integral.

Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vectors, $\{\boldsymbol{X}_1, \boldsymbol{X}_2, \ldots \}$, by low discrepancy sequences.  This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.

This tutorial describes  low discrepancy sequences  and their quality measures.  We demonstrate the performance gains possible with quasi-Monte Carlo methods.  Moreover, we describe how to formulate problems to realize the most increase in performance using quasi-Monte Carlo methods.  We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean.

@@

%\bibliographystyle{apalike} \bibliography{FJH23,FJHown23}
